Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-09-20 elsevier_test/elsevier_test.Rmd
many many changes

---
title: Food-web complexity flattens the fitness landscape of an insect herbivore
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich, Switzerland
    footnote: Corresponding Author
  - name: Jordi Bascompte
    email: jordi.bascompte@ieu.uzh.ch
    affiliation: Another University
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
abstract: 
  Studies of natural selection and fitness landscapes usually treat the network of interacting species as a “black box”. Given that the loss of biodiversity is simplifying the structure of ecological networks, there is a pressing need to answer the question: how does network complexity affect natural selection and the fitness landscape of associated species? To answer this question, we conducted a field experiment that manipulated the complexity of a food web associated with a galling insect herbivore. To maintain complex food webs, we allowed the entire community of natural enemies to attack insect galls on 64 plants in a common garden setting. To create simple food webs, we excluded a guild of three larval parasitoids by bagging galls on 64 different plants; therefore, mortality in this treatment was primarily due to a single egg parasitoid that attacks prior to gall formation. We then measured herbivore survival as a function of three key gall traits in each treatment. We found that more traits were under selection in the simple vs. complex food web. This occurred because different parasitoid species impose different selection pressures on gall traits, thereby minimizing relative fitness differences among insect galls with different phenotypes. Our work suggests that more complex food webs allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution to environmental change. Species-interaction networks provide a mechanistic link between community and population ecology, as they describe who interacts with whom in a community. Similarly, fitness landscapes provide a mechanistic link between population and evolutionary ecology, as they describe the influence of natural selection on phenotypic variation. It remains unclear, however, how the network structure of species interactions shapes the fitness landscape. Understanding this relationship is likely key for developing a predictive ecology across scales of biological organization. To examine the relationship between network structure and the fitness landscape, we conducted a field experiment that manipulated the network of trophic interactions (simple vs. complex) influencing the fitness of an insect herbivore. We then quantified the fitness landscape of herbivores in each treatment by measuring herbivore survival as a function of multiple phenotypic traits. We found that more traits were under selection in the simple vs. complex trophic network. This occurred because different natural enemies impose different selection pressures on herbivore traits, thereby minimizing relative fitness differences among herbivore phenotypes in the complex network. Our work suggests that more complex trophic networks allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution of populations to environmental change.

@@

_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_

@@

## Load required libraries ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
library(visreg)
library(viridis) # for color palette

@@

  # subset data for analysis
  filter(phenology == "early", Location == "tree",
         platy > 0 | ectos > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  mutate(gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg_parasitoid = as.numeric(ifelse(platy > 0, 1, 0)))

# convert treatment to factor to use as 'by' variable in generalized additive models
gall_selection.df <- gall_selection.df %>%
  mutate(Treatment_contr.sum=factor(Treatment.focus), 
         sc.gall_size = as.numeric(scale(Gall_Height_mm)),
         sc.clutch_size = as.numeric(scale(gall_individuals)),
         sc.female_preference = as.numeric(scale(Density_per_100_shoots)))
contrasts(gall_selection.df$Treatment_contr.sum) <- 'contr.sum'

## Create separate datasets for quantifying selection gradients ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))

## Custom functions to avoid repetitive code ----
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Genotype/Plant_Position/Gall_Number), 
        family = binomial(link = logit))
}

gam.plot <- function(gam.model, ...) {
  plot(gam.model, se = TRUE, seWithMean = T, shift = mean(predict(gam.model)), trans = function(x) {exp(x)/(1+exp(x))}, ...)
}

# experimenting with boot.case gave bootstrapped distributions that weren't centered near the model mean estimate.
# also, using 'refit.smooth = T' also generated some very skewed distributions
# the recipe below provided the most sensible estimates for my data
gradient.calc <- function(mod, phenotype, covariates = NULL) {
  gam.gradients(mod, phenotype, covariates, parallel = 'multicore', ncpus = 32, 
                refit.smooth = F,
                n.boot = 1000,
                se.method = 'boot.para', #'posterior', #'boot.para', #'permute', 'boot.case'#  
                standardized = T)
}

insect_FL <- function(mod, phenotype, covariates = NULL) {
  fitness.landscape(mod, phenotype, covariates, parallel = 'multicore', ncpus = 32,
                    PI.method = 'boot.para',
                    refit.smooth = F,
                    PI.interval = c(0.025,0.975),
                    plt.density = 100)
}

get_insect_FL.df <- function(FL){
  data.frame(mean_fitness = FL$Wbar, FL$points, t(FL$WbarPI)) %>%
    mutate(lower_2.5 = X2.5., upper_97.5 = X97.5.)
}

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

inverse_logit <- function(x) exp(x)/(1+exp(x))

#Beta_avg_grad <- function(logistic_model, mean_fitness) {
#  brackets <- predict(logistic_model, type = "response") * (1 - predict(logistic_model, type="response"))
#  gradient <- mean(brackets) * fixef(logistic_model) / mean_fitness 
#  return(gradient)
#}

## Calculate directional selection gradients
# following method of Janzen and Stern 1998, Evolution
Beta_avg_grad <- function(.) {
  mean_fitness <- inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

## Bootstrap fitness estimates
bootstrap_fitness <- function(logistic_model, newdata, bootstraps, intervals=c(0.025,0.975)){
  model_matrix <- model.matrix(~ sc.gall_size + sc.clutch_size + sc.female_preference, newdata)

@@

  # Organize data
  return(list(absolute_fitness = absolute_fitness_df, relative_fitness = relative_fitness_df))
}
```

> *There is a much more insidious kind of extinction: the extinction of ecological interactions.* 
> *(Janzen, D.H. 1974. The deflowering of Central America. Natural History. 83:48–53).*

>>>>

---
title: Food-web complexity alters the fitness landscape of an insect herbivore
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: a,b
    footnote: Corresponding Author
  - name: Christopher J. Greyson-Gaito
    affiliation: a,c
  - name: Arezoo Sootodeh
    affiliation: a
  - name: Brendan Locke
    affiliation: d
  - name: Jordi Bascompte
    affiliation: b
address:
  - code: a
    address: University of British Columbia, Department of Zoology, 6270 University Blvd., Vancouver, BC, V6T 1Z4, Canada
  - code: b
    address: University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, Zurich, 8057, Switzerland
  - code: c
    address: University of Guelph, Department of Integrative Biology, 50 Stone Rd. East, Guelph, ONT, N1G 2W1, Canada
  - code: d
    address: Humboldt State University, Department of Biological Sciences, 1 Harpst St., Arcata, CA, 95521, USA
abstract: "Studies of natural selection and fitness landscapes usually treat the network of interacting species as a “black box”. Given that     the loss of biodiversity is simplifying the structure of ecological networks, there is a pressing need to answer the question: how does network complexity affect natural selection and the fitness landscape of associated species? To answer this question, we conducted a field experiment that manipulated the complexity of a food web associated with a galling insect herbivore. To maintain complex food webs, we allowed the entire community of natural enemies to attack insect galls on 64 plants in a common garden setting. To create simple food webs, we excluded a guild of three larval parasitoids by bagging galls on 64 different plants; therefore, mortality in this treatment was primarily due to a single egg parasitoid that attacks prior to gall formation. We then measured herbivore survival as a function of three key gall traits in each treatment. We found that more traits were under selection in the simple vs. complex food web. This occurred because different parasitoid species impose different selection pressures on gall traits, thereby minimizing relative fitness differences among insect galls with different phenotypes. Our work suggests that more complex food webs allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution to environmental change."

@@

<!-- Species-interaction networks provide a mechanistic link between community and population ecology, as they describe who interacts with whom in a community. Similarly, fitness landscapes provide a mechanistic link between population and evolutionary ecology, as they describe the influence of natural selection on phenotypic variation. It remains unclear, however, how the network structure of species interactions shapes the fitness landscape. Understanding this relationship is likely key for developing a predictive ecology across scales of biological organization. To examine the relationship between network structure and the fitness landscape, we conducted a field experiment that manipulated the network of trophic interactions (simple vs. complex) influencing the fitness of an insect herbivore. We then quantified the fitness landscape of herbivores in each treatment by measuring herbivore survival as a function of multiple phenotypic traits. We found that more traits were under selection in the simple vs. complex trophic network. This occurred because different natural enemies impose different selection pressures on herbivore traits, thereby minimizing relative fitness differences among herbivore phenotypes in the complex network. Our work suggests that more complex trophic networks allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution of populations to environmental change.-->

@@

## Load required libraries ----
library(tidyverse)
library(cowplot) # pretty default ggplots
#library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
#library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
#library(visreg)
library(viridis) # for color palette
library(lme4) # for generalized linear mixed models

@@

  # subset data for analysis
  filter(phenology == "early", Location == "tree",
         platy > 0 | ectos > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  mutate(gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg_parasitoid = as.numeric(ifelse(platy > 0, 1, 0)),
         sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
         sc.clutch_size = as.numeric(scale(gall_individuals)), 
         sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
         sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
         sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))

## Create separate datasets for quantifying selection gradients ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") #%>%
  #mutate(sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
  #       sc.clutch_size = as.numeric(scale(gall_individuals)), 
  #       sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
  #       sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
  #       sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion", ectos < 1) #%>%  # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  #mutate(sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
  #       sc.clutch_size = as.numeric(scale(gall_individuals)), 
  #       sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
  #       sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
  #       sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))

mean(filter(gall_selection.df, Treatment.focus == "Ectoparasitoid exclusion")$ectos) # less than 3% of larva were parasitized by an ectoparasitoid in the exclusion experiment.

## Custom functions to avoid repetitive code ----

## Test for overdispersion
overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

## Transform logit to probability
inverse_logit <- function(x) exp(x)/(1+exp(x))

## Calculate selection gradients following method of Janzen and Stern 1998, Evolution
Beta_avg_grad <- function(.) {
  mean_fitness <- mean(predict(., type = "response")) #inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

## Calculate intensity/opportunity for selection
intensity_of_selection <- function(.) {
  mean_fitness <- mean(predict(., type = "response")) # inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  relative_fitness <- predict(., type = "response") / mean_fitness 
  var(relative_fitness)
}

## Bootstrap fitness estimates
bootstrap_fitness <- function(logistic_model, fixed_effects, newdata, bootstraps, intervals=c(0.025,0.975)){
  model_matrix <- model.matrix(formula(fixed_effects), newdata)

@@

  # Organize data
  return(list(absolute_fitness = absolute_fitness_df, relative_fitness = relative_fitness_df))
}

## Number of bootstrap replicates
n_boots_analysis <- 1000
n_boots_plots <- 100

```

AIC comparison suggests gam is better than quadratic approximation; however, I'm worried that the GAM model is overfitting. I feel like the quadratic approximation, while not providing the best fit, will still preven the model from overfitting.

Now the question is: normal glm vs. mixed-model? Maybe I need to answer, why not the mixed model?

- There is no theory developed for them. 
- Relative fitness occurs at different hierarchical levels instead of the level of the sampled population.
- Less straightforward.

Mixed-models are also good for modeling heterogeneity, but if I'm interested in a general population level, then maybe I don't have to worry about that. 
```{r GAM, include=F}
control_gam <- gam(gall_survival ~ s(sc.gall_size) + s(sc.log.clutch_size, k=9) + s(sc.log1p.female_preference), control_df, family=binomial(link=logit))
plot(control_gam, trans = inverse_logit)

treatment_gam <- gam(gall_survival ~ s(sc.gall_size) + s(sc.log.clutch_size, k=9) + s(sc.log1p.female_preference), treatment_df, family=binomial(link=logit))
plot(treatment_gam, trans = inverse_logit)

control_gam_quad <- gam(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2), control_df, family=binomial(link=logit))

treatment_gam_quad <- gam(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2), treatment_df, family=binomial(link=logit))

AIC(control_gam, control_gam_quad)
AIC(treatment_gam, treatment_gam_quad)

hist(gall_selection.df$sc.gall_size)
hist(gall_selection.df$sc.log.clutch_size)
hist(gall_selection.df$sc.log1p.female_preference)

control_glm <- glm(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2) + sc.gall_size:sc.log.clutch_size + sc.gall_size:sc.log1p.female_preference + sc.log.clutch_size:sc.log1p.female_preference, control_df, family=quasibinomial())#family=binomial(link=logit))
summary(control_glm)
visreg::visreg(control_glm, scale = "response")

treatment_glm <- glm(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2) + sc.gall_size:sc.log.clutch_size + sc.gall_size:sc.log1p.female_preference + sc.log.clutch_size:sc.log1p.female_preference, treatment_df, family=quasibinomial())#family=binomial(link=logit))
summary(treatment_glm)

visreg::visreg(control_glm, scale = "response")
visreg::visreg(treatment_glm, scale = "response")

test_glm <- glm(gall_survival ~ Treatment.focus*(sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2) + sc.gall_size:sc.log.clutch_size + sc.gall_size:sc.log1p.female_preference + sc.log.clutch_size:sc.log1p.female_preference), gall_selection.df, family=quasibinomial())#family=binomial(link=logit))
summary(test_glm)

visreg::visreg2d(test_glm, xvar="sc.gall_size", yvar="sc.log.clutch_size", scale="response")
```

```{r, Cryptic effects of parasitism on gall morphology, include=F}
# how does this analysis compare to Heath?
# does it make sense to look for the interaction?

control_gall_size <- lmer(sc.gall_size ~ gall_survival +
                           (1|Genotype:Plant_Position:Gall_Number),
                         data = control_df)
summary(control_gall_size)
visreg::visreg(control_gall_size, xvar="gall_survival")

piecewiseSEM::rsquared(control_gall_size)

flip.control_gall_size <- lmer(gall_survival ~ sc.gall_size +
                           (1|Genotype:Plant_Position:Gall_Number),
                         data = control_df)
summary(flip.control_gall_size)
visreg::visreg(flip.control_gall_size, xvar="sc.gall_size")

lm.control_gall_size <- lm(Gall_Height_mm ~ gall_survival,
                         data = control_df)
summary(lm.control_gall_size)
visreg::visreg(lm.control_gall_size, xvar="gall_survival")

lmer.control_gall_size <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype:Plant_Position:Gall_Number),
                         data = control_df)
summary(lmer.control_gall_size)
visreg::visreg(lmer.control_gall_size, xvar="gall_survival")

lmer.control_gall_size_2 <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype/Plant_Position/Gall_Number),
                         data = control_df)
summary(lmer.control_gall_size_2)
visreg::visreg(lmer.control_gall_size_2, xvar="gall_survival")

piecewiseSEM::rsquared(flip.control_gall_size)

treatment_gall_size <- lmer(sc.gall_size ~ gall_survival +
                           (1|Genotype:Plant_Position:Gall_Number),
                         data = treatment_df)
summary(treatment_gall_size)
visreg::visreg(treatment_gall_size, xvar="gall_survival")

lm.treatment_gall_size <- lm(Gall_Height_mm ~ gall_survival,
                         data = treatment_df)
summary(lm.treatment_gall_size)
visreg::visreg(lm.treatment_gall_size, xvar="gall_survival")

lmer.treatment_gall_size <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype:Plant_Position:Gall_Number),
                         data = treatment_df)
summary(lmer.treatment_gall_size)
visreg::visreg(lmer.treatment_gall_size, xvar="gall_survival")

lmer.treatment_gall_size_2 <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype/Plant_Position/Gall_Number),
                         data = treatment_df)
summary(lmer.treatment_gall_size_2)
visreg::visreg(lmer.treatment_gall_size_2, xvar="gall_survival")

piecewiseSEM::rsquared(lmer.treatment_gall_size_2)

lmer.treatment_sc.gall_size_2 <- lmer(sc.gall_size ~ gall_survival + sc.clutch_size + sc.female_preference + (1|Genotype/Plant_Position/Gall_Number),
                         data = treatment_df)
summary(lmer.treatment_sc.gall_size_2)

lm.treatment_sc.gall_size <- lm(sc.gall_size ~ gall_survival,
                         data = treatment_df)
summary(lm.treatment_sc.gall_size)

summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2), data=treatment_df))
summary(lmer(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2) + (1|Genotype/Plant_Position/Gall_Number), data=treatment_df))

summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2) , data=control_df))
summary(lmer(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2) + (1|Genotype/Plant_Position/Gall_Number), data=control_df))

pair_control <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_control$diff)
length(pair_control$diff)

t.test(pair_control$`1`, pair_control$`0`, paired = T)

pair_treatment <- treatment_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_treatment$diff)
length(pair_treatment$diff)
pair_treatment

t.test(pair_treatment$`1`, pair_treatment$`0`, paired = T)

pair_control_ectos <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(ectos)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_control_ectos$diff)
length(pair_control_ectos$diff)

t.test(pair_control_ectos$`1`, pair_control_ectos$`0`, paired = T)

pair_control_platy <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(egg_parasitoid)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_control_platy$diff)
length(pair_control_platy$diff)

t.test(pair_control_platy$`1`, pair_control_platy$`0`, paired = T)

gsg::moments.differentials(z=control_df$sc.gall_size, W = control_df$gall_survival)
gsg::moments.differentials(z=treatment_df$sc.gall_size, W = treatment_df$gall_survival)

summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size, control_df))
summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size, treatment_df))

test_control_df <- bind_rows(mutate(control_df, type = 0), mutate(filter(control_df, gall_survival == 1), type = 1))
summary(lm(sc.gall_size ~ type, test_control_df)) # virtually the same as above

test_treatment_df <- bind_rows(mutate(treatment_df, type = 0), mutate(filter(treatment_df, gall_survival == 1), type = 1))
summary(lm(sc.gall_size ~ type, test_treatment_df)) # virtually the same as above

# can also look at it this way and just look at the mean difference before and after selection
test_control_df %>%
  group_by(type) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

test_treatment_df %>%
  group_by(type) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

# does this enable me to estimate the confounding effect of parasitism on gall size? 
# Note that the intercept is no longer centered on zero...
summary(lmer(sc.gall_size ~ type + (1|Genotype/Plant_Position/Gall_Number), test_control_df))
summary(lmer(sc.gall_size ~ type + (1|Genotype/Plant_Position/Gall_Number), test_treatment_df))

moments_control_df <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

gsg::moments.differentials(z=moments_control_df$mean_gall_size, W=moments_control_df$gall_survival) # does this mean anything?

moments_treatment_df <- treatment_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

gsg::moments.differentials(z=moments_treatment_df$mean_gall_size, W=moments_treatment_df$gall_survival)

### MOST PROMISING ANALYSIS THUS FAR THAT MAY ENABLE ME TO ASSESS THE BIAS OF PARASITISM ON OBSERVED SELECTION GRADIENTS.
test_again_control <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_again_control$diff_gall_size)
t.test(test_again_control$survive_mean_gall_size, test_again_control$all_mean_gall_size, paired = T)

test_ectos_control <- control_df %>%
  filter(pupa > 0 | ectos > 0) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_ectos_control$diff_gall_size)
t.test(test_ectos_control$survive_mean_gall_size, test_ectos_control$all_mean_gall_size, paired = T)

test_platy_control <- control_df %>%
  filter(pupa > 0 | egg_parasitoid > 0) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_platy_control$diff_gall_size)
t.test(test_platy_control$survive_mean_gall_size, test_platy_control$all_mean_gall_size, paired = T)

test_again_treatment <- treatment_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_again_treatment$diff_gall_size)
t.test(test_again_treatment$survive_mean_gall_size, test_again_treatment$all_mean_gall_size, paired = T)
```

```{r Calculate Mean and SD of Gall Traits at Different Scales, include=F, eval=F}

## Calculate mean and SD of gall traits at different scales ----
summary.individual_level <- gall_selection.df %>%
  group_by(Treatment.focus) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean, sd), na.rm = TRUE) %>%
  mutate(Level = "Individual")

mean.gall_size.control <- filter(summary.individual_level, Treatment.focus == "Control")$Gall_Height_mm_mean
sd.gall_size.control <- filter(summary.individual_level, Treatment.focus == "Control")$Gall_Height_mm_sd
mean.gall_size.treatment <- filter(summary.individual_level, Treatment.focus == "Ectoparasitoid exclusion")$Gall_Height_mm_mean
sd.gall_size.treatment <- filter(summary.individual_level, Treatment.focus == "Ectoparasitoid exclusion")$Gall_Height_mm_sd

summary.gall_level <- gall_selection.df %>%
  group_by(Treatment.focus, Gall_Number) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean), na.rm = TRUE) %>%
  group_by(Treatment.focus) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean, sd), na.rm = TRUE) %>%
  mutate(Level = "Gall")

mean.clutch_size.control <- filter(summary.gall_level, Treatment.focus == "Control")$gall_individuals_mean
sd.clutch_size.control <- filter(summary.gall_level, Treatment.focus == "Control")$gall_individuals_sd
mean.clutch_size.treatment <- filter(summary.gall_level, Treatment.focus == "Ectoparasitoid exclusion")$gall_individuals_mean
sd.clutch_size.treatment <- filter(summary.gall_level, Treatment.focus == "Ectoparasitoid exclusion")$gall_individuals_sd

summary.plant_level <- gall_selection.df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean), na.rm = TRUE) %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean), na.rm = TRUE) %>%
  group_by(Treatment.focus) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean, sd), na.rm = TRUE) %>%
  mutate(Level = "Plant")
